---
title: "Reproductible Reaseatch_Project1"
author: "Serge NYOKA"
date: "23 octobre 2016"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=8, warning=FALSE, messsage=FALSE)
```
    
Reproductible Research : Course Project 1

## 1. Loading and preprocessing the data

```{r}
#Reading data file in the csv format.
dt_activity <- read.csv("activity.csv")
str(dt_activity)
```


## 2. What is mean total number of steps taken per day ?

```{r}
#Summing the steps for each day. 60 observations for 60 days 
steps_daily <- aggregate(steps ~ date, dt_activity, sum) 
head(steps_daily)
```

Plotting Histogram

```{r echo=FALSE}
hist(steps_daily$steps, main = "Total number of steps per day", xlab = "Number of steps", col = 5)
```

#The mean is equal to 10766 and the median to  10765
summary(steps_daily)


## 3. What is the average daily activity pattern ?


```{r}
#Calculating the mean and the median of the total numberof steps per day
steps_mean <- mean(steps_daily$steps)
steps_median <- median(steps_daily$steps)

#For all days and for each interval we calculate the average steps
steps_interval <- aggregate(steps ~ interval, dt_activity, mean)
```

Plotting the average number of steps per day for each interval

```{r echo=FALSE}
plot(steps_interval$interval, steps_interval$steps, type = "l", xlab = "Interval", ylab = "Average steps number", main = "Average of each day steps's number by interval", col = "blue")
```

```{r}
#Compute the maximum average steps of intervals
interval_maximum <- steps_interval[which.max(steps_interval$steps), 1]
interval_maximum 
```

## 4. Imputing missing values

```{r}
#Calculating and reporting the total number of missing value
missing_data_number <- sum(!complete.cases(dt_activity))
missing_data_number

dt_activity_imput <- transform(dt_activity, steps = ifelse(is.na(dt_activity$steps), steps_interval$steps[match(dt_activity$interval, steps_interval$interval)], dt_activity$steps))
dt_activity_imput[as.character(dt_activity_imput$date) == "2012-10-01", 1] <- 0

#Compute total steps per day imputed
steps_daily_imput <- aggregate(steps ~ date, dt_activity_imput, sum)
```
```{r}
head(dt_activity_imput)
```
Visualizing the imputed data without all NAs


#Histogram of total number steps for each day

```{r echo=FALSE}
hist(steps_daily_imput$steps, xlab = "Number of steps", col = "gray", main = "Total number of steps per day")
```

```{r}
#Calculate and report the mean and median total number of steps taken per day
steps_mean_imput <- mean(steps_daily_imput$steps)
steps_mean_imput

steps_median_imput <- median(steps_daily_imput$steps)
steps_median_imput

#Difference between mean, median and total number steps of the 2 dataset (imputed and non-imputed)
delta_mean <- steps_mean_imput - steps_mean
delta_mean

delta_median <- steps_median_imput - steps_median
delta_median

delta_steps <- sum(steps_daily_imput$steps) - sum(steps_daily$steps)
delta_steps
```

## 5. Are there differences in actyivity patterns between weekdays and weekends ?

Differences %in% activity patterns between weekdays and weekends
```{r}
dayPos <- weekdays(as.Date(dt_activity_imput$date))
for(i in 1:length(dayPos)) {
  if(dayPos[i] %in% c("samedi", "dimanche")) dayPos[i] = "weekend"
  else dayPos[i] = "weekday"
}

dt_activity_imput$dayPosition <- as.factor(dayPos)

steps_interval_imput <- aggregate(steps ~ interval + dayPosition, dt_activity_imput, mean)

library(lattice)
```

Plotting ...
  
```{r echo=FALSE}  
xyplot(steps_interval_imput$steps ~ steps_interval_imput$interval|steps_interval_imput$dayPosition, xlab = "Interval", ylab = "steps", main = "Average steps for each day by interval", layout = c(1,2), type = "l")
```
The difference is visible on these 2 charts 